### 2025-01-23 paper/main.tex
Shuffle parts of the Discussion.

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{bm}
\usepackage{caption}
\usepackage{subcaption}
% \captionsetup[subfigure]{justification=justified,singlelinecheck=false}
\usepackage{tikz}
\usepackage{newtxmath} % MER: Prettier math
\usepackage{cleveref} % MER: Allows \Cref (but use Section X.Y not Subsection X.Y)
\usepackage{booktabs} % MER: Prettier tables

\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{float}
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{placeins}
\usetikzlibrary{quotes}
\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,positioning,fit,petri}

\graphicspath{{figures/}}

\captionsetup[subfigure]{justification=justified,singlelinecheck=false}

\title{In-silico molecular enrichment and clearance of the human intracranial space}
\author[1]{Marius Causemann}
\author[1]{Miroslav Kuchta}
\author[2]{Rami Masri}
\author[1,3,*]{Marie E. Rognes }
\affil[1]{Department of Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway}
\affil[2]{Brown University, Providence, Rhode Island, USA}
\affil[3]{K. G. Jebsen Centre for Brain Fluid Research}
\affil[*]{meg@simula.no}

\newcommand{\rami}[1]{\textcolor{blue}{#1}}
\newcommand{\mer}[1]{\textcolor{magenta}{#1}}
\newcommand{\mar}[1]{\textcolor{violet}{#1}}
\newcommand{\mk}[1]{\textcolor{orange}{#1}}
\newcommand{\discuss}[1]{\textcolor{red}{#1}}
\newcommand{\fixme}[1]{\textcolor{red}{#1}}
\newcommand{\draft}[1]{\textcolor{lightgray}{#1}}
%\newcommand{\brain}{\Omega_{\rm{PAR}}} 
%\newcommand{\sas}{\Omega_{\rm CSF}}
%\newcommand{\spinal}{\Gamma_{\rm SAS}}
%\newcommand{\skull}{\Gamma_{\rm{skull}}}
%\newcommand{\gin}{g_{\rm{influx}}}

\newcommand{\R}{\mathbb{R}}
\newcommand{\foralls}{\forall \,}

%\affil[+]{these authors contributed equally to this work}

%\keywords{Keyword1, Keyword2, Keyword3}

%% The problem we are addressing here is that we would like to be able to
%% understand, integrate and quantify the role of different mechanisms -
%% including effects associated with individual variations - on molecular
%% enrichment and clearance of human intracranial spaces. And we think of
%% this as a step on the way to predict tracer enrichment after
%% intrathecal injection in a personalized medicine context. Which is
%% important because of drug delivery/treatment in cancer, diagnostics
%% for CSF and dementia disorders, and clearance in neurodegenerative
%% disease.

\begin{abstract}
  The mechanisms of intracranial solute transport are fundamental to
  human brain health and function, with alterations often linked to
  disease and functional impairment, and with distinct opportunities
  for personalized diagnostics and treatment. However, our
  understanding of these mechanisms and their interplay remains
  incomplete, in part due to the complexity of integrating insights
  across scales, between species and from different modalities. Here,
  we combine geometric model reduction and mixed-dimensional
  modelling, multi-modal magnetic resonance images, and high
  performance computing to construct and explore a high-fidelity
  in-silico model of human intracranial molecular enrichment and
  clearance. This model predicts the temporal and spatial spreading of
  a solute within an image-derived geometric representation of the
  subarachnoid space, ventricular system and brain parenchyma,
  including branching networks of surface perivascular spaces
  (PVSs). Our findings highlight the significant impact of
  cerebrospinal fluid (CSF) production and intracranial pulsatility on
  molecular enrichment following intrathecal tracer injection. We
  demonstrate that low-frequency vasomotion induces moderate CSF flow
  rates in surface PVS networks that substantially enhance tracer
  enrichment, and that impaired enrichment is a natural direct
  consequence of enlarged PVSs. This openly available technology
  platform thus provides an opportunity for integrating separate
  observations on diffusion in neuropil, vascular dynamics,
  intracranial pulsatility, CSF production, and efflux, and for
  exploring questions relating to drug delivery and clearance in the
  human brain environment.
\end{abstract}
\begin{document}

\flushbottom
\maketitle
% * <john.hammersley@gmail.com> 2015-02-09T12:07:31.197Z:
%
%  Click the title above to edit the author information and abstract
%
\thispagestyle{empty}

%\mer{MER: Target length: 5000 words (Intro, Results, Discussion) plus <3000 words Methods: 700 words Introduction. 3100 words Results. 1200 words Discussion.}

\iffalse
\paragraph*{Abbreviations}
\begin{itemize}
\item
  Subarachnoid space (SAS)
\item
  Cerebrospinal fluid (CSF)
\item
  Perivascular space (PVS), perivascular spaces (PVSs)
\end{itemize}
\fi

\section*{Introduction}

% What is the big picture problem and why is it important? 
The mechanisms underlying molecular transport within the intracranial
space are fundamental to human brain health and
function. Neurodegenerative diseases such as Alzheimer's, Parkinson's,
and Huntington's disease are all associated with abnormal accumulation
of protein aggregates together with alterations in transport and
clearance characteristics~\cite{rasmussen2018glymphatic,
  harrison2020impaired, eide2023plasma, liu2024glymphatic}. Moreover,
sleep and conversely sleep-deprivation play a definite yet enigmatic
role in modulating molecular enrichment and
clearance~\cite{xie2013sleep, eide2021sleep, eide2022altered,
  miao2024brain, hauglund2025norepinephrine}. In the last decade,
established theories have been challenged by new findings revealing a
greater degree of molecular movement and
exchange~\cite{iliff2012paravascular, ringstad2017glymphatic,
  louveau2017understanding, proulx2021cerebrospinal,
  bohr2022glymphatic}, as well as substantial variability in
enrichment and clearance between individuals and between patient
cohorts~\cite{ringstad2018brain, eide2021direction, eide2021impaired,
  eide2022altered}. These observations provide distinct opportunities
(and challenges) for personalized medicine e.g.~for tailored
intrathecal delivery of chemotherapy~\cite{lohela2022glymphatic} and
for early diagnostics of impaired brain
clearance~\cite{eide2021clinical, van2024human}. In spite of their
importance, our understanding of these mechanisms is incomplete with
open challenges and significant debate -- in part relating to the
translation of knowledge between scales, species, experimental
protocols, clinical cohorts, and individuals.

Perivascular pathways along the brain surface and within the brain
parenchyma have long been hypothesized to serve a designated role in
this context~\cite{rennels1985evidence, zhang1990interrelationships,
  ichimura1991distribution, carare2008solutes, iliff2012paravascular,
  foley2012realtime, hannocks2018molecular, van2024caa}. Recently, Eide and
Ringstad~\cite{eide2024functional} and Yamamoto et
al~\cite{yamamoto2024perivascular} demonstrated that perivascular
spaces (PVSs) define preferential pathways for molecular transport in
humans, with delayed periarterial enrichment in dementia
subtypes~\cite{eide2024functional}. Perivascular flow of cerebrospinal
fluid (CSF) clearly contributes to this transport, and is inherently
associated with vascular pulsations~\cite{hadaczek2006perivascular,
  iliff2013cerebral, bedussi2018paravascular, mestre2018flow,
  boster2023artificial, hirschler2024region}. Fluid mechanics
considerations point at intracranial pressure differences and shorter
wavelength vascular wall pulsations as drivers of directional net flow
and convection in the PVS~\cite{bilston2003arterial, rey2018pulsatile,
  daversin2020mechanisms, kedarasetti2020functional,
  gjerde2023directional, nozaleda2024arterial}, while longer waves
such as the pulse wave primarily contribute to oscillatory flow and
dispersion~\cite{asgari2016glymphatic, sharp2019dispersion, thomas2019fluid,
  kedarasetti2020arterial, troyetsky2021dispersion, 
  martinac2021phase}. In the bigger picture, CSF is produced by the
choroid plexus~\cite{damkier2013cerebrospinal,
  steffensen2018cotransporter, liu2020direct}, pulsates through the
ventricular system, cisterns, and subarachnoid space (SAS) in
synchrony with cardiac, respiratory, and neural
waves~\cite{greitz1993pulsatile, wagshul2011pulsating,
  sweetman2011cerebrospinal, fultz2019coupled, vinje2019respiratory,
  eide2021direction, causemann2022human, williams2023neural,
  zimmermann2023total}, and drains via the dural sinuses, meningeal
lymphatics, cranial nerves, or other efflux
pathways~\cite{proulx2021cerebrospinal}. However, how these
physiological factors and physical mechanisms integrate to enhance or
impair human intracranial molecular transport over larger spatial
scales and longer time scales remain unknown. A related key question
is to what extent surface PVSs are separated from the SAS by
structural barriers~\cite{zhang1990interrelationships,
  weller2005microscopic, bedussi2017paravascular,
  pizzo2018intrathecal, mestre2022periarteriolar,
  mollgard2023mesothelium, smets2024perivascular, eide2024functional},
and in turn to what extent such structural compartmentalization is a
prerequisite for effective perivascular transport.

% What are our research questions and how do we answer them
In this study, by leveraging geometric model reduction and
mixed-dimensional modelling~\cite{masri2024modelling}, structural
magnetic resonance (MR) images~\cite{hodneland2019new}, and high
performance computing, we introduce an integrated computational model
of intracranial molecular enrichment and clearance. Focusing on the
interplay between perivascular pathways, pulsatility and CSF flow
dynamics, the model predicts the temporal evolution and spatial
distribution of a solute concentration within a detailed geometric
representation of the human SAS and ventricular system, networks of
surface PVSs and the brain parenchyma. In terms of transport dynamics,
we account for heterogeneous diffusion, dispersive mixing induced by
cardiac and respiratory pulsatility in the CSF spaces and PVSs,
convective fluid flow driven by CSF production and peristaltic
pumping, as well as solute exchange and clearance across
semi-permeable membranes. Comparing with glymphatic MRI
studies~\cite{ringstad2017glymphatic, ringstad2018brain,
  watts2019measuring}, the in-silico predictions accurately represent
molecular enrichment patterns, timing and intercompartmental
distributions. This open platform~\cite{ZENODO} thus provides a
technological opportunity for qualitatively and quantitatively
exploring key open questions relating to molecular movement within the
human brain environment such as the role of pulsatility, perivascular
pathways, structural compartmentalization, or morphological
pathologies.

% Main findings and outlook
By exploring this high-dimensional parameter space, we propose that
the balance between CSF production and intracranial pulsatility is key
to shaping the large-scale features of intracranial enrichment
patterns, with the potential to span a wide range of individual and
cohort variability. Moreover, we predict that CSF production, cardiac-
and respiratory pulsatility is not sufficient to explain early
perivascular enrichment, but that fluid flow induced by low-frequency
vasomotion in surface periarterial spaces (on the order of 13 \textmu
m/s) is sufficient, even in the absence of a structural
compartmentalization of the PVS. Conversely, enlarged PVSs in the SAS
will cause a substantial reduction in cardiac- and vasomotion-driven
flow velocities, strongly delay perivascular transport, and thus
impair intracranial enrichment. These findings transfer, reconcile,
integrate and extend insights from clinical, experimental, and
theoretical studies, and lay a new foundation for in-silico studies of
personalized intrathecal drug delivery and brain clearance.

%% \begin{itemize}
%% \item
%%   Need better understanding of transport of solutes in intracranial spaces.
%% \item
%%   Need to integrate separate observations e.g. on diffusion in
%%   neuropil, PVS flow, net flow, effect of dispersion
%% \item
%%  Need to transfer and integrate insights from animal models,
%%   computational models and theoretical insights into the human case. 
%% \item
%%  Controversies need to be resolved.
%% \end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
%% \mer{MER: Target ~3100 words in 6 subsections}

%% \mer{MER: Results #1 is about 500 words}
\subsection*{In-silico predictions of intracranial molecular enrichment and clearance after intrathecal injection}
\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{figures/figure1.png}
     \caption{A) Illustration of the spatial domains of the model: CSF-filled space (ventricles and SAS) $\Omega_{\rm CSF}$, parenchyma $\Omega_{\rm PAR}$ and the PVS surrounding arteries $\Lambda_{\rm artery}$ and the veins $\Lambda_{\rm vein}$ and their interfaces and boundaries (see Methods); B) 3D rendering of the actual geometry used in this study parenchyma, CSF (clipped, blue), arterial network (red), venous network (blue)); C) arterial network (red) and venous network (blue); D) illustration of the modelled clinical tracer injection procedure; E) time of simulated first tracer arrival after intrathecal injection (concentration exceeds $0.1\,$mol/l); F) time of first tracer arrival in the arterial PVS (top) and venous PVS (bottom); G) sagittal, coronal and axial view of computed tracer concentrations over T1-image 1, 6, 12 and 24\,h after injection (low concentrations transparent, pial surface in pink, arachnoid membrane in cyan, and arteries in dark red); H) total amount of tracer in each cranial compartment over the first 24\,h; I) average tracer concentration in each compartment over the first 24\,h. }
     \label{fig:results1}
\end{figure}
Using previously published multi-modal magnetic resonance imaging
(MRI) data~\cite{hodneland2019new, deistung2009tof,
  schweser2012quantitative, reichenbach2012future,
  deistung2017overview}, we construct a multiscale computational
representation of the human intracranial compartments consisting of
the CSF spaces and brain parenchyma as three-dimensional (3D) domains
and with the PVSs surrounding major surface arteries and veins as
embedded networks of topologically one-dimensional (1D) curves
(\Cref{fig:results1}A--C). We consider a solute concentration field,
varying in space and time, in the 3D domains and in the PVS networks,
and assume that the solute can cross between these compartments
through semi-permeable membranes. As the drivers and modes of
intracranial transport are under substantial
debate\cite{smith2019going, proulx2021cerebrospinal,
  bohr2022glymphatic, hladky2022glymphatic, betsholtz2024advances},
our first target is to establish a baseline model accounting for a
reasonably conservative set of mechanisms and their integrated effect
over a timescale of several minutes to a few days. To this end, we
assume that the solute will (i) diffuse within all compartments, with
diffusivity depending on the effective properties of the relevant
medium\cite{sykova2008diffusion}; (ii) experience significant
dispersive effects due to the pulsatile flow of CSF induced by the
cardiac and respiratory cycles\cite{vinje2019respiratory,
  sharp2019dispersion, ray2021quantitative, troyetsky2021dispersion};
and (iii) be convected by a (small) net flow of CSF resulting from
production in the choroid plexus with CSF efflux across the upper
convexity\cite{hornkjol2022csf} and from the peristaltic pumping
effect of pulse wave pulsations in surface periarterial
spaces~\cite{mestre2018flow, gjerde2023directional}
(\Cref{fig:results1}D). Mathematically, this model is represented by a
mixed-dimensional system of coupled time-dependent partial
differential equations\cite{masri2024modelling}, which we solve
numerically with high accuracy using a mass-conserving finite element
scheme and the FEniCS finite element software\cite{alnaes2015fenics,
  kuchta2020assembly} (see Methods). The computational framework and
associated software are all openly available~\cite{ZENODO}.

Simulating a glymphatic MRI protocol~\cite{ringstad2017glymphatic,
  ringstad2018brain, eide2024functional}, we then predict the
spreading of $0.5$ mmol intrathecally injected Gadobutrol after its
appearance at the craniocervical junction
(\Cref{fig:results1}E--G). After one hour, the in-silico
tracer moves upwards in the SAS frontally of the brainstem, and
quickly reaches the supratentorial regions. Here, it spreads both
posterior through the quadrigeminal cistern and the longitudinal
fissure, and anteriorly through the outer SAS, reaching the top of the
cerebral cortex after around 12 hours. After 24 hours, the tracer
covers most of the brain surface (with the exception of some posterior
regions) and has penetrated several centimeters into the parenchymal
tissue. This pattern is reflected in the mean tracer concentrations in
each compartment, where the CSF space reaches its peak of
$1.4\,$mmol/l after 2 hours, followed by the arterial PVS
concentration peak ($2.9\,$mmol/l after 4 hours)
(\Cref{fig:results1}I). While the mean concentrations in these
compartments drop soon after peaking, the parenchymal tissue
slowly enriches with tracer over the first 24 hours, with a final
value of $0.3$ mmol/l. Overall, about 40\% of the total amount of
tracer remains in the cranium 24 hours post-injection, with the largest
share in the CSF (53\%), followed by the parenchyma (35\%), the
venous PVS (8\%) and the arterial PVS
(5\%) (\Cref{fig:results1}H).

% \mer{MER: Results #2 is about 600 words}
\subsection*{Reduced CSF pulsatility strongly shifts intracranial enrichment patterns}
\begin{figure}%[h!]
\centering 
\includegraphics[width = 0.95\textwidth]{figures/figure2.png}
\caption{\textbf{The balance between CSF pulsatility and production shapes molecular enrichment patterns.}
A) CSF spaces: the third and fourth ventricle (V3\&4), lateral ventricles (LV) and SAS; 
B) Boundaries and interfaces: the spinal subarachnoid space (SSAS), upper and lower arachnoid membrane (AM-U and AM-L), pia (Pia) and lateral ventricles (LV); 
C) Schematic of the CSF flow model; 
D) CSF flow induced by CSF production with outflow allowed across the
upper convexity (AM-U); 
}
\label{fig:csf}
\end{figure}
\begin{figure}
\ContinuedFloat
\caption{(\textbf{cont.})
E) Histograms of the production-induced CSF velocities in the SAS and third and fourth ventricle (V3 \& V4) in terms of the relative frequency of the mean flow speed in each computational cell weighted by its volume; 
F) CSF flow induced at peak systolic blood inflow; 
G) Cardiac-induced dispersion factor $R_c$;
H) CSF flow induced at peak respiratory expansion; 
I) Respiration-induced dispersion factor $R_r$, $D = (1 + R_c + R_r) D^{\rm Gad}$; J--M): Histograms of the cardiac-induced flow velocities (J),
cardiac-induced dispersion factors (K),
respiratory flow velocities (L), respiratory dispersion factor (M);
N--Q): sagittal, coronal and axial planes of tracer concentration after 12h for different model variations: baseline (N), no CSF production (O), low dispersion (P), and high dispersion (Q).
}
\end{figure}

The enrichment patterns observed clinically after intrathecal
injection of contrast differ substantially between subjects and
between pathological conditions~\cite{ringstad2018brain,
  eide2021direction, eide2021impaired, eide2022altered}. These
neurological conditions are also associated with alterations in the
pulsatile flow of CSF in the ventricular system and
SAS~\cite{eide2021direction}. Clearly, key physiological factors such
as cardiac pulsatility and respiration easily differ between
individuals and between both pathological and physiological states. We
therefore next asked to what extent -- and how -- variations in CSF
pulsatility would affect the intracranial enrichment characteristics.

For the CSF dynamics in our baseline model, we combine the net
contribution to CSF flow induced by CSF production with the integrated
dispersive effects of cardiac and respiratory pulsatile CSF flow
(\Cref{fig:csf}A--C). Imposing a constant CSF production of
$400\,$ml/day across the surface of the lateral ventricles, while
allowing for efflux across the upper convexity, yields a total CSF
pressure drop of $26\,$mPa (0.00020 mmHg) (\Cref{fig:csf}D) with a
maximum flow velocity of $1.85\,$mm/s (in the aqueduct) and a mean
velocity in the SAS of $2.6$\textmu m/s (\Cref{fig:csf}E). To examine the
dispersive effects of pulsatile CSF flow, we employ the theory of
shear-augmented dispersion together with computational fluid dynamics
to determine a dispersion factor $R$ enhancing the effective solute
diffusivity (see Methods). For the cardiac contribution, we compute
CSF pressure and flow fields at peak systolic blood inflow,
corresponding to a reduction in CSF space volume at a total rate of
$6\,$ml/s~\cite{baledent2014imaging, causemann2022human} in the SAS
and $0.31\,$ml/s across the lateral ventricle
surface\cite{vinje2019respiratory}. This scenario sets up a pressure
drop of $10\,$Pa ($0.075$ mmHg) between the lateral ventricles and the
spinal SAS and a maximum flow velocity of $19.8\,$cm/s
(\Cref{fig:csf}F, J). Assuming a cardiac frequency of $1\,$Hz, we
infer that this cardiac-induced pulsatile CSF flow increases the
effective diffusion by more than two orders of magnitude in the
aqueduct and near the cisterna magna, but has little effect ($R <
1$) in most of the SAS (\Cref{fig:csf}G, K). For the respiratory
contribution, we employ the same methodology, but with a total rate of
$1\,$ml/s \cite{gutierrez2022effect} in the SAS and $0.121\,$ml/s
\cite{liu2024using} in the lateral ventricles, yielding a respiratory
peak flow volume of $1.121\,$ ml/s at the craniocervical
junction. While the resulting flow velocities are only about one
fourth of their cardiac-induced counterparts (see \Cref{fig:csf}H, L),
respiratory dispersive mixing reaches a factor of up to 320 due the
lower respiratory frequency of $0.25\,$Hz (\Cref{fig:csf}I, M).

Now, to examine how CSF pulsatility affects the enrichment patterns,
we consider three variations of the baseline: (i) no CSF production,
(ii) reduced pulsatility and thus decreased dispersion ($0.1 \times
R$), and (iii) higher pulsatility with increased dispersion ($10
\times R$). Without CSF production, transport is considerably delayed
(\Cref{fig:csf}N, O). Even after $12$ hours, tracer remains in the
subtentorial regions around the cerebellum, the brain stem and in the
surrounding CSF. Interestingly, the lack of CSF production instead
allows tracer to travel upwards through the ventricular system,
reaching the third ventricle after around $12$ hours. On the other
hand, if the CSF pulsatility is reduced, we observe rapid transport
towards the upper convexity of the cranium, as in the baseline model,
but the tracer spreads exclusively within the anterior regions
(\Cref{fig:csf}P). This feature can be attributed to the CSF flow
bifurcation posterior to the ambient cistern (\Cref{fig:csf}D):
without sufficient diffusion, the tracer is unable to cross into the
posterior SAS. Indeed, with higher dispersion, the tracer moves
through the quadrigeminal cistern and further upwards into the
longitudinal fissure, with also enrichment of the cerebellum
(\Cref{fig:csf}Q).

% \mer{MER: Results #3 is about 800 words}
\subsection*{Perivascular flow shapes and accelerates molecular enrichment}
\label{sec:pvs_flow_results}

The PVSs are recognized across species as critical pathways for solute
transport in and around the brain\cite{}, and thus as potential
targets for enhancing brain drug delivery and metabolic waste
clearance\cite{}. However, whether CSF flows more rapidly in PVSs and
what the forces and mechanisms required to drive such flow are, remain
as key points of debate~\cite{bohr2022glymphatic,
  van2024caa}. Motivated by experimental observations in animal
models~\cite{iliff2012paravascular, iliff2013cerebral, mestre2018flow,
  bedussi2018paravascular}, there is now a remarkable body of
literature on modelling perivascular fluid flow and
transport~\cite{bilston2003arterial, asgari2016glymphatic,
  rey2018pulsatile, daversin2020mechanisms, sharp2019dispersion,
  thomas2019fluid, kedarasetti2020functional, kedarasetti2020arterial,
  troyetsky2021dispersion, martinac2021phase, gjerde2023directional,
  nozaleda2024arterial}. Here, we ask how these proposed mechanisms
would translate from idealized geometries to human vascular networks
and moreover evaluate their integrated effect in the context of
intracranial solute transport.

Our periarterial network extends from the internal carotid arteries
and basilar artery through up to 18 bifurcations to reach upstream
network ends located within the SAS or up to 6 mm inside the
parenchyma (\Cref{fig:results1}A--C, \Cref{fig:pvs}A) and includes
major surface arteries of radius 0.5--1.4 mm. Imposing the pressure
field induced by CSF production at the periarterial network ends
induces ultraslow steady CSF flow of variable direction in these PVSs,
with an average velocity of 0.08 \textmu m/s (antegrade) and a maximum
velocity of 0.55 \textmu m/s (\Cref{fig:pvs}B). On the other hand,
traveling waves of arterial wall motion (\Cref{fig:pvs}C), such as the
pulse wave or other vasomotion~\cite{vanveluw2020vasomotion,
  munting2023spontaneous, bojarskaite2023sleep, broggini2024long,hauglund2025norepinephrine},
also induce net directional flow in the
PVS~\cite{kedarasetti2020arterial, kedarasetti2020functional,
  coenen2021lubrication, gjerde2023directional, nozaleda2024arterial}
-- of magnitude and direction depending on the amplitude, frequency,
and length of the waves and the characteristics of the perivascular
network~\cite{gjerde2023directional}. Applying a
semi-analytic model of the net flow induced by peristalsis in
perivascular networks~\cite{gjerde2023directional} (see Methods), we
estimate that the cardiac pulse wave alone, traveling at a frequency
of 1 Hz with a wavelength of 2.0 m and a 1\% wall
displacement~\cite{jung2021novel}, will induce mainly antegrade PVS
flow with an average net velocity of 0.92 \textmu m/s while reaching up
to 7.31 \textmu m/s near larger, ventral vessels such as the MCA
(\Cref{fig:pvs}D). The same theory predicts that strong ultraslow
vasomotions, if traveling antegrade at 0.1Hz with a wavelength of 0.02
m and a 10\% displacement~\cite{broggini2024long}, will induce both
retrograde and antegrade net PVS flow with an average velocity of
13.05 \textmu m/s and maximum velocity 54.44 \textmu m/s
(\Cref{fig:pvs}E).

We study the effect of such rapid PVS flow on intracranial molecular
enrichment by comparing the more conservative baseline model and a
high PVS flow model, where the former still includes the net flow
contributions from CSF production and the cardiac pulse wave, while
the latter additionally includes the ultraslow vasomotion
contribution. In both models, tracers were first observed at the basal
artery after 48 minutes (\Cref{fig:pvs}F), defined by their first-time
arrival (FTA), the time at which the concentration first exceeds $0.1$
mmol/l. At all upstream locations along the middle cerebral arteries
(MCAs) and anterior cerebral artery (ACA), we found substantially
reduced FTAs with higher PVS flow (\Cref{fig:pvs}G), up to 2.5 hours
earlier in the left M2 segment of the MCA. PVS concentrations peaked
before the surrounding tissue (0:48h vs 1:12h at the MCA and 1:00h vs
1:48h at MCA2) with high PVS flow, while these peaks occurred nearly
simultaneously in the baseline model. The earlier appearance and the
delay in time-to-peak between the PVS and surrounding tissue clearly
indicates the directionality of tracer enrichment -- it first arrives
in the PVS and subsequently spreads into the tissue and SAS
(\Cref{fig:pvs}H, I), and especially so with higher PVS flow. These
observations also hold on an aggregated level: computing the total
amount of tracer in the PVS as a function of the distance to the
arterial network roots, we find accelerated tracer transport with
higher PVS flow, especially at earlier time points (2--9 hours)
(\Cref{fig:pvs}J). Moreover, differences in tracer transport along the
PVS translate to altered enrichment patterns on the whole-organ
scale. For instance, after 4-6 hours, the faster-moving tracer in the
PVS is clearly visible in the adjacent spaces to the ACA, MCA, and
other arteries (\Cref{fig:pvs}L). \mer{One more sentence on effect of
  high PVS flow on tracer enrichment outside the PVS.} In contrast,
there are no clear signs of early enrichment surrounding the PVS in
the baseline model (\Cref{fig:pvs}K).

A natural question is whether early PVS enrichment could be the result
of increased dispersion rather than net flow in the
PVS\cite{asgari2016glymphatic,sharp2019dispersion,bojarskaite2023sleep,asgari2016glymphatic,troyetsky2021dispersion}. Interestingly,
even increasing the dispersion factor by $100 \times$ in the PVS,
induced only minor changes in the global spreading rate compared to the
baseline (\Cref{fig:pvs}G, M), thus indicating a negative answer to
this question.
\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{figures/figure3.png}
    \caption{\textbf{Perivascular flow shapes and accelerates molecular enrichment.}}
    \label{fig:pvs}
\end{figure}
\begin{figure}
\ContinuedFloat
  \caption{\textbf{(cont.)}
    A) Surface arterial network colored by arterial radius with vessel locations labeled;
    B) Net PVS flow induced by CSF production (cf.~\Cref{fig:csf}D) -- histograms show the relative frequencies of local periarterial velocities. Positive values indicate antegrade PVS flow, while negative values indicate retrograde PVS flow;
    C) Illustration of the concept of peristaltic pumping;
    D) Estimated net PVS flow induced by pulse wave peristaltic pumping;
    E) Estimated net PVS flow induced by vasomotion/slow wave peristaltic pumping;
    F) Comparison of mean tracer concentration in the PVS and over the outer PVS surface at key locations of the arterial tree, for the baseline (lower) and high PVS flow (upper) models, with labels: ``peak'': time-to-peak (h), ``$\Delta t$'': time difference between peaks (h), and ``FTA'': first-time tracer arrival (h).
    G) first-time arrival for the main trunks of the periarterial network over the distance from the closest network root node (BA, ICA-R or ICA-L for the baseline, the high PVS flow and the high PVS dispersion models);
    H, I) 2D slices showing zoom-in on the region surrounding the MCA-R and MCA2-R for the baseline (upper) and high PVS flow (lower) models. Innermost circle indicates the artery (blue), the surrounding annulus is the PVS (color represents concentration value), the surroundings are parenchyma and/or CSF spaces, separated by the pia (blue: towards the parenchyma, pink: towards the CSF));
    J) Total amount of tracer in the periarterial spaces as a function of the distance to the nearest network root over time for the baseline (left) and the high PVS flow (right) models;
    K--M) sagittal (upper) and coronal (lower) views of the simulated tracer concentrations overlayed on the T1-weighted MR image after 4 and 6h for the baseline (K), high PVS flow (L) and high PVS dispersion (M) models (concentrations \discuss{below threshold masked}, pial surface outlined in pink, arachnoid membrane in cyan, and arteries in dark red).}
\end{figure}

% \mer{MER: Results #4 is about 500 words}
\subsection*{Structural versus functional compartmentalization of perivascular spaces}

Human and rodent observations indicate that tracers concentrate in perivascular spaces surrounding the pial and subarachnoid vasculature \cite{zhang1990interrelationships,zhang1992directional, bedussi2017paravascular, mestre2018flow, bedussi2018paravascular, eide2024functional}. However, it remains unclear whether such enrichment patterns necessitate a structural barrier, such as a membrane with limited permeability, or if the patterns could result from enhanced flow or mixing in these areas alone. We therefore next investigate how the permeability of the interface between the PVS and the surrounding CSF affects tracer enrichment around the major arterial trunks (\Cref{fig:compartmentalization}A--B). To this end, we compare models with high and low  permeabilities, representing a highly permeable PVS-CSF interface (functional compartmentalization) and a less permeable interface (structural compartmentalization), respectively. For the low permeability model, we set the permeability to $3.8 \cdot 10^{-7}\,$m/s, consistent with previous estimates for the endfoot sheath surrounding penetrating arterioles \cite{koch2023estimates}, which we consider to be a lower bound for the surface PVS-CSF permeability. For the high permeability model, we increase this permeability by a factor of 100. For both scenarios, we consider the high PVS flow regime as examined in the previous section.

At the basal artery (BA), tracer appears within one hour in the surrounding CSF in both models (\Cref{fig:compartmentalization}C, E). With a more permeable PVS-CSF interface, tracer quickly crosses from the CSF into the PVS, resulting in similar peak concentrations of 25\,mmol/l in the PVS and at its outer surface, while values up to 50\,mmol/l are attained in the vicinity. In contrast, the low permeability model exhibits substantially lower perivascular tracer enrichment, with a maximum concentration of 10\,mmol/l after approximately two hours.
%The high permeability model allows for more tracer crossing into the PVS, leading to higher concentrations further downstream. Notably, crossover into the parenchyma is also observed, most clearly at 4 hours.
For the middle cerebral arteries (MCA-R and MCA-L), tracer appears
first in the PVS in both models (\Cref{fig:compartmentalization}D, E),
though with higher concentrations in the high permeability model (up
to 24 mmol/l). We also observe that the higher permeability allows the
tracer to leak out of the PVS and spread within the Sylvian fissure to
a greater extent. \discuss{With a small delay, more tracer appears via
  the CSF pathway (\Cref{fig:compartmentalization}F, peak time
  difference).} At the anterior cerebral artery (ACA-A2), the first tracer arrival in the PVS occurs after 3 hours, with a peak at 5 hours. The concentration peak in the PVS is tailed by a higher peak in the CSF (\Cref{fig:compartmentalization}E), indicating tracer arrival through a second pathway; i.e., directly through the CSF-filled space outside of the PVS. Similar patterns are observed for ACA-A3.

Summarizing and quantifying these observations
(\Cref{fig:compartmentalization}F), we find that neither the time of
first arrival nor the time-to-peak are substantially affected by the
PVS-CSF permeability at any of the periarterial segments considered
(BA, MCAs, ACAs). However, the concentration difference between the
PVS and its surrounding CSF quickly decreases with higher
permeability, as does the time lag between the concentration peaks in
these domains. Thus, fast transport along the PVS is not contingent
upon a structural compartmentalization of the PVS, whereas a sharp
concentration gradient between PVS and CSF is unlikely without a
restricting barrier.
\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{figures/figure4.png}
    \caption{\textbf{Does early perivascular enrichment rely on a structural compartmentalization of the PVS?}
    A) The PVS-CSF membrane permeability regulates the exchange of solutes between the PVS and surrounding CSF spaces; 
    B) The arterial network with highlighted regions of interest; 
    C) 2D slices showing zoom-in to the region surrounding the basal artery after 1, 2, 3, and 4 hours for a low (above) and high (below) PVS-CSF permeability model. Innermost circle indicates the artery (blue), the surrounding annulus is the PVS (color represents concentration value), the surroundings are parenchyma and/or CSF spaces, separated by the pia (blue: towards the parenchyma, pink: towards the CSF));
    D) As in (C) for regions surrounding (from upper to lower) the ACA-A3, ACA-A2, MCA-R and MCA-L segments;}
    \label{fig:compartmentalization}
\end{figure}
\begin{figure}
\ContinuedFloat
\caption{(\textbf{cont.})
  E) Comparison of mean tracer concentration in the PVS (dashed) and over the
outer PVS surface (dotted) for the BA, MCA-R, MCA-L, ACA-2, and ACA3 segments over time for the high (purple) and low permeability (yellow) models;
  F) Effect of varying the PVS-CSF permeability on first-time-of-arrival (FTA), time difference ($\Delta t$) between concentration peaks in the PVS and surrounding tissue, time of PVS peak concentration, PVS peak concentration and difference in peak concentration between PVS and surrounding tissue.}
\end{figure}

% \mer{MER: Target Results 5 300 words}
\subsection*{Enlarged PVSs delay periarterial and intracranial molecular enrichment}

Enlarged PVSs are associated with impaired periarterial tracer
transport and more generally with cognitive decline as well as
certain subtypes of
dementia~\cite{bown2022physiology,eide2024functional}. A key
question is thus whether enlarged PVSs alone leads to delayed perivascular
and intracranial tracer enrichment. To address this question from a
fluid and transport dynamics perspective, we ask our in-silico model
to predict the integrated effect of perivascular dilation on PVS flow
velocities and tracer enrichment patterns (\Cref{fig:enlargement}A).

Again we compute CSF production-, cardiac-, and vasomotion-induced net PVS flow velocities, but now in a perivascular geometry with an increased outer radius ($R^{\rm dilated}_2 = 3 R_1$) compared to the previous ($R^{\rm control}_2 = 2 R_1$), which corresponds to a $2.67 \times$ increase in PVS cross-section area.
%% \begin{align*}
%%   A_1 = \pi (R_2^2 - R_1^2) = \pi (2^2 R_1^2 - R_1^2) = 3 \pi R_1^2 \\
%%   A_2 = \pi (\hat{R}_2^2 - R_1^2) = \pi (3^2 R_1^2 - R_1^2) = 8 \pi R_1^2 \\
%%   A_2 / A_1 = 8/3 = 2.76
%% \end{align*}
The enlargement yields notable and non-uniform changes in the flow
dynamics. On the one hand, the pressure-induced flow velocities
increase from a mean of 0.08 \textmu m/s to 0.33 \textmu m/s due to
the reduced effective resistance of the PVS. However, the peristaltic
pumping becomes considerably less effective: the cardiac and
vasomotion-induced net CSF flow velocities drop from a mean of 0.92
\textmu m/s to 0.17 \textmu m/s and from 13.05 \textmu m/s to 2.33
\textmu m/s, respectively (\Cref{fig:enlargement}B--D). In total, the
net CSF velocity reduces, which in turn alters the tracer enrichment
within the dilated PVS. Notably, we observe later tracer arrival in
all MCA segments, with up to \discuss{120} min later arrival in the
MCA2s for the high PVS flow scenario (\Cref{fig:enlargement}E, F). For
the baseline model, which omits the vasomotion-induced net PVS flow,
the effect is less pronounced in the MCA but clearly persists and is
evident in the MCA2s (\Cref{fig:enlargement}F). Finally, we note that
the larger volume of the dilated PVS results in a greater total
accumulation of tracer in the PVS from 3--24 hours and reduced
enrichment of the parenchyma at 6h, while the mean \discuss{PVS}
concentration drops slightly (\Cref{fig:enlargement}G,H).
\begin{figure}
    \centering
    \includegraphics[width=1 \textwidth]{figures/figure5.png}
    \caption{\textbf{PVS enlargement reduces net PVS flow and delays tracer transport}
    A) Modelling dilated PVS by extending the outer PVS boundaries;
    B) mean and max PVS flow velocities with normal/control and dilated PVS;
    C--D) 3D rendering of vasomotion-induced PVS flow in the normal (C) and dilated (D) periarterial networks; 
    E) 2D slices showing zoom-in to the region surrounding the MCA-R after 1, 2, 3, and 4\,h with normal (lower) and dilated (upper) PVS. Innermost circle indicates the artery (blue), the surrounding annulus is the PVS (color represents concentration value), the surroundings are parenchyma and/or CSF spaces, separated by the pia (blue: towards the parenchyma, pink: towards the CSF)); 
    F) First-time arrival for different branches of the arterial tree for the baseline model (with pressure-driven and cardiac-induced net PVS flow) and the high PVS flow model (also including vasomotion-induced net PVS flow) for normal and dilated PVS;
    G) total tracer content in the CSF, parenchyma and arterial and venous PVS at 1,3,6,12 and 24\,h for the four models from F;
    H) mean concentration in the CSF, parenchyma and arterial and venous PVS over time for the four models from F.
    }
    \label{fig:enlargement}
\end{figure}

% \mer{MER: Target Results 6 400 words}
%\subsection*{The perivenous contribution to molecular clearance of CSF tracers}

%\begin{itemize}
%\item Look at clearance phase 12h-24h-48h-72h.
%\item Look at concentration in the veins
%\item Look at concentration in the parenchyma
%\item Consider low-pressure sink for the perivenous spaces.
%\end{itemize}

%% \subsection*{Clearance of intra-parenchymal solutes}

%% \begin{itemize}
%% \item
%%   Clearance of a molecular concentration in the parenchyma, different
%%   molecular sizes - ranging from Gadubutrol to amyloid-$\beta$/$\tau$.
%% \end{itemize}

\section*{Discussion}

%% Summary of findings
We have presented a high-fidelity in-silico model of molecular
enrichment and clearance in human intracranial spaces, enabling
tailored predictions of the influence of CSF space and (peri)vascular
morphology, physiological factors such as cardiac and respiratory
pulsatility, as well as of pathological conditions such as enlarged
PVSs. A key observation is that the balance between CSF production and
intracranial pulsatility significantly shapes the large-scale features
of molecular enrichment, with e.g.~ventricular tracer reflux after
intrathecal injection in the absence of CSF production, and preference
towards anterior brain regions with reduced pulsatility. Indeed, the
simulated distribution patterns cover a substantial range of the
individual variations observed clinically \cite{ringstad2018brain}. In
terms of perivascular pathways, we find that even moderate CSF flow on
the order of 0.1--1.0 \textmu m/s in the surface PVSs results in
earlier tracer enrichment around major cerebral arteries. This effect
is more pronounced -- with up to a three-hour difference in time of
arrival -- with PVS flow rates of approximately 10 \textmu m/s. Our
models of CSF flow in perivascular networks, based on first principles
and asymptotic analysis, predict net flow of such magnitudes and both
antegrade and retrograde flow within the perivascular network,
occurring with or without pressure differences at the network
ends. The PVS may thus function as highways facilitating rapid
transport, even in the absence of structural barriers, though a sharp
concentration difference between the PVS and surrounding tissues is
unlikely without such barriers. Finally, dilated PVS will cause a
substantial reduction in cardiac- and vasomotion-driven flow
velocities, which in turn impedes perivascular transport.

%% Comparison with others

% Tracer enrichment predictions
While molecular enrichment of the brain and surrounding CSF spaces
have been extensively studied in animal models, especially in murine
models in connection with studies of the glymphatic system, there are
fewer reports of contrast enrichment in human subjects over 0--24h. We
therefore mainly compare our in-silico predictions with the series of
papers by Eide and colleagues~\cite{ringstad2017glymphatic,
  ringstad2018brain, eide2021sleep, eide2024functional} and Watts et
al~\cite{watts2019measuring}. Ringstad et al.~\cite{ringstad2018brain}
report of tracer enrichment in a centripetal pattern, primarily in
regions near large cerebral surface arteries, peaking in the CSF
spaces between 6 and 24 hours, with parenchymal tracer content still
increasing after 24 hours, but with large individual
differences. Watts et al~\cite{watts2019measuring} report of a similar
enrichment pattern and a concentration peak in the SAS of $\approx$0.5
mmol/L occuring after 10 to 15 hours. Our in-silico enrichment
patterns agree with these observations, though with an earlier
\discuss{$\sim$5 + plus spinal transit time)} rather than later peak
in the CSF spaces. In particular, we note that the substantial
individual variation observed clinically is comparable to the
variation in in-silico enrichment patterns associated with reduced CSF
production or reduced intracranial pulsatility. \mer{One sentence on
  tracer content comparison.} In terms of perivascular tracer
enrichment, Eide and Ringstad~\cite{eide2024functional} report of
first-time tracer appearance in the PVS of the MCA after 37.8 $\pm$
47.0 (M1), 53.1 $\pm$ 50.5 (M2) and 82.6 $\pm$ 61.5 (M3) min, thus
with just a 15.3 min delay between M2 and M1. Comparing with the our
FTAs in the MCA segments, we note that the baseline model predicts a
more-than 3 hour difference in FTA between the M1 and M2. Instead, the
high PVS flow scenario predicts a 24 min delay between these
segments. Moreover, the high PVS flow scenario predicts FTAs of
84--108 min (M1), 84--146 min (M2) which overlaps with the previous
ranges \discuss{modulo spinal transit time}. We interpret these
results as indicative of a net CSF flow in human surface PVS beyond
what would be induced by CSF production and cardiac peristalsis alone.
% M2 - M1 (mean) ~ 15.3 min

% PVS flow predictions. 
The existence, magnitude, and directionality of flow in surface and
parenchymal PVSs have been the subject of active debate for
decades~\cite{rennels1985evidence, bilston2003arterial,
  hadaczek2006perivascular, carare2008solutes, iliff2012paravascular,
  iliff2013cerebral, bakker2016lymphatic, mestre2018flow,
  bedussi2018paravascular, rey2018pulsatile, thomas2019fluid,
  daversin2020mechanisms, kedarasetti2020functional,
  martinac2021phase, vanveluw2020vasomotion, bohr2022glymphatic,
  boster2023artificial, gjerde2023directional,
  nozaleda2024arterial}. In mice, Mestre et al~\cite{mestre2018flow}
observe CSF flow in surface perivascular regions surrounding branches
of the MCA with average flow speeds of 18.7 \textmu m/s and net flow
in the antegrade direction, but also regions with retrograde
flow. Similarly, Bedussi et al.~\cite{bedussi2018paravascular} report
of oscillatory flow, with an average (net) CSF velocity of 17 $\pm$ 2
\textmu m/s in the antegrade direction. Using physics-informed neural
networks, Boster et al~\cite{boster2023artificial} estimate PVS
velocities of $12.75 \pm 6.25$ \textmu m/s. Less is known about the
magnitude and directionality of CSF flow in surface PVSs in
humans. Here, assuming that human surface PVS follow the major surface
vessels and are CSF-filled but otherwise open regions of width
proportional to the vessel radius, our estimates indicate that the CSF
pressure differences due to CSF production, cardiac peristaltic
pumping and low-frequency vasomotion all yield spatially-varying net
CSF flow with both antegrade and retrograde PVS segments. This
observation is thus in agreement with the experimental
observations~\cite{mestre2018flow, bedussi2018paravascular}, and
inherently supports the original notion that antegrade and retrograde
solute transport along PVSs may coexist~\cite{rennels1985evidence}. In
terms of magnitude, the contribution from CSF production and cardiac
wall motion are small (0.08 \textmu m/s and 0.92 \textmu m/s on
average), but intriguingly, our estimates of the contribution from
vasomotion (13.05 \textmu m/s) are comparable to the experimental
reports. We do note that this estimate is based on the presence of
rhythmic waves of vasomotion at a frequency of 0.1Hz, a wavelength of
20 mm, and a wall displacement of 10\% -- in the antegrade direction,
all values currently associated with significant
uncertainty~\cite{gokina1996electrical, vanveluw2020vasomotion,
  broggini2024long}.

% Compartmentalization of the PVS
To what extent are surface PVSs in communication with the surrounding
SAS and to what extent are these structurally separated compartments?
Originally, Weller and coauthors~\cite{zhang1990interrelationships,
  zhang1992directional, weller2005microscopic} identified thin sheaths
of pial cells surrounding human surface arteries and penetrating
arterioles (but not veins or venules). On the other hand, Bakker and
coauthors report that the subarachnoid space, the cisterns, ventricles
and penetrating periarteriolar spaces form a continuous CSF
space~\cite{bedussi2017paravascular}. Thorne and colleagues study the
architectures of rat perivascular spaces in detail, emphasizing the
presence of openings (stomata) on the interface between the
vasculature and the CSF within the SAS~\cite{pizzo2018intrathecal,
  hannocks2018molecular}. Further, Mestre et
al~\cite{mestre2022periarteriolar} study the properties of pial
perivascular spaces in mice, and report of pial cells forming sheaths
for larger surface arteries and partially cover smaller surface
arteries, with higher coverage in ventral SAS regions. Our results
demonstrate that the PVSs may act as rapid transport pathways even
with only a partial barrier between the PVS and surrounding CSF, thus
contributing to quantifying the properties of the human PVS-CSF
interface~\cite{eide2024functional}.

% Variations with sleep, exercise
Lifestyle factors such as sleep~\cite{xie2013sleep, miao2024brain,
  bojarskaite2023sleep, eide2021sleep, vinje2023human,
  hauglund2025norepinephrine, larsen2024sleep},
exercise~\cite{holstein2018voluntary, yildiz2022immediate}, and
alcohol intake~\cite{lundgaard2018beneficial} affect molecular
enrichment and clearance of the brain. In particular, pial arteries
display higher-amplitude low-frequency vasomotion during NREM and
intermediate sleep states, while REM sleep is associated with reduced
PVS width~\cite{bojarskaite2023sleep}. Peristaltic pumping is more
effective under both of these configurations, resulting in higher
estimated net PVS flow velocities~\cite{gjerde2023directional}. Even
higher PVS velocities would amplify the PVS flow effects simulated
here, with earlier arrival times along the PVSs, clear enrichment in
adjacent spaces, and accelerated molecular enrichment
overall. Moreover, sleep is linked with low-frequency oscillations in
human CSF flow~\cite{fultz2019coupled}. As we have shown here, the
level of dispersion in the CSF spaces strongly shapes molecular
enrichment patterns. While we, at baseline, account for the dispersion
induced by cardiac and respiratory pulsatility, the high dispersion
scenarios considered can be viewed as representative of additional
pulsatility induced by e.g.~neural waves~\cite{fultz2019coupled,
  williams2023neural} or other respiration patterns depending on
activity level. More broadly, we consider this as a key strength of
the current in-silico framework: allowing for integrating and
exploring the mechanistic macroscale implications of changes in
pulsatility at smaller spatial or temporal scales.

% Variations with disease states
%cohort variations, individual variations, physiological factor
%variations. Enlarged PVS reduce transport at the whole cerebral scale.

In terms of limitations, our primary constraint modelling-wise is the
geometric representation of the vascular (and therefore also
perivascular) networks. Concretely, we account for no vessels below
the original MRI resolution limits, and thus only a part of the
cerebral surface vasculature and effectively not the cortical
vessels. In addition, and again due to lack of resolution, there is
uncertainty associated with the location of the vessel segments
relative to the CSF spaces and parenchyma, with some vessels
surrounded by both domains and some vessels completely entering the
parenchyma before returning to the SAS. Higher resolution MRI data
(T1w and e.g. time-of-flight, QSM or other modalities) in more
subjects would immediately allow for more detailed studies of this
aspect. As a result of the lack of an accurate representation of the
cortical and subcortical white matter vasculature, we have considered
a conservative transport model within the parenchyma, modelling
extracellular diffusion alone. To compensate, we have focused on
reporting detailed predictions in the CSF spaces and surface PVS after
not more than 24 hours. In particular, we note that, we observe tracer
accumulation at the perivascular network ends after \fixme{24} hours,
similar to that previously observed for
microspheres~\cite{bedussi2018paravascular, mestre2018flow}, which
here can be interpreted as an artifact of the incomplete network
creating up a barrier for tracer movement. Previously, we have found
that enrichment within the parenchyma is well-represented by $\approx
3\times$ enhanced diffusion augmented by local clearance e.g.~across
the blood-brain barrier or diffusion-convection with a complex pattern
of spatially-varying convective velocities of 1-10 \textmu
m/min~\cite{vinje2023human}. We will incorporate such transport
mechanisms within the parenchyma in future work. Finally, we also note
that we simulate the intracranial distribution of a tracer
concentration appearing at the craniocervical junction rather than as
an intrathecal injection. We therefore overestimate the amount of
tracer entering into the intracranial spaces; however, due to the
linearity of the mathematical model, it is valid to directly interpret
the in-silico concentrations relative to the total amount of tracer.

\section*{Methods}

\subsection*{Intracranial compartments: ventricular system, SAS, and brain parenchyma}

From T1-weighted MR images of a human subject~\cite{hodneland2019new},
we first automatically segment the brain parenchyma and CSF spaces,
including the ventricular system and SAS, using
Synthseg~\cite{billot2023robust,billot2023synthseg}
(\Cref{fig:results1}A--B). We next manually adjust the segmentation to
accurately represent the connections between CSF spaces (aqueduct,
median aperture, tentorium cerebelli), smoothen, and finally extract
surface representations of the outer (arachnoid) boundary, the pial
membrane and other interfaces. Conforming to these surface and
interface representations, we generate a tetrahedral mesh $\Omega$
representing the complete intracranial volume as the union of the
parenchyma $\Omega_{\rm PAR}$ and CSF spaces $\Omega_{\rm CSF}$, using
fTetWild~\cite{hu2020fast} (\Cref{fig:csf}A--B). The resulting
computational mesh consists of $233\,592$ vertices and $1\,290\,131$
mesh cells, which vary between $0.22$ mm and $8.9$ mm in mesh cell
diameter. The parenchyma has a total volume of $1318$ ml, whereas the
ventricles and the outer SAS contribute $32$ ml and $329$ ml,
respectively, to a total CSF volume of $361$ml, in agreement with
recent estimates~\cite{hladky2024regulation}.

%% In addition, we slightly modify the segmentation to ensure correct
%% connectivity of the ventricular system. In particular, we elongate
%% the aqueduct and enforce no communication between the ventricles
%% and the surrounding CSF spaces except for the median aperture.
%% After morphological smoothing of the segmented image, we extract
%% surface representations of the outer SAS boundary, the pial
%% membrane and the components of the ventricular system.

%% Note that the widely reported value of $150$\,ml of intracranial
%% CSF volume is likely an underestimate. Recent estimates indicate
%% that a value of $200$ to $300$\,ml is more accurate (see
%% \cite{hladky2024regulation} and references therein

\subsection*{Periarterial and perivenous spaces}

To represent surface networks of periarterial and perivenous spaces,
we use Kiminaro~\cite{william_silversmith_2021_5539913} to separately
skeletonize time-of-flight angiography (ToF) and quantitative
susceptibility mapping (QSM) images from the same human
subject~\cite{hodneland2019new}. This technique yields networks of
one-dimensional curves, each curve $\Lambda^i$ indicating the
centerline of a blood vessel segment and labeled with its lumen radius
$R_1^i$ (\Cref{fig:results1}C, \Cref{fig:pvs}A). We also assign an
outer radius $R_2^i > R_1^i$ to each segment $i$, form the annular
cylinder ensheathing $\Lambda^i$ and define the union of these as the
PVS. The periarterial network graph consists of $12\,708$ edges
connected at $12\,562$ inner nodes and with $147$ end nodes, and the
associated domain is denoted by $\Lambda_a$. We identify three of the
end nodes -- corresponding to the two internal carotid arteries (ICAs)
and the basilar artery (BA) -- as root nodes, and designate the other
ends as leaf nodes. The resulting perivenous network graph consists of
$\fixme{N}$ edges connected at $\fixme{N}$ inner nodes and with
$\fixme{N}$ end nodes, and the associated domain is denoted by
$\Lambda_v$. \discuss{Finally, we label the outer surface of the PVSs
  associated with $\Lambda_a$ and $\Lambda_v$, by $\Gamma_a$ and
  $\Gamma_v$ respectively.}

%% The resulting vascular networks are further smoothened using a
%% spline interpolation technique, and represented as graphs with
%% nodes connected by edges.

\subsection*{Molecular transport equations}

We model diffusion, convection, and exchange of a molecular solute in
the PVS networks $\Lambda_a$ and $\Lambda_v$, in the CSF spaces
$\Omega_{\rm CSF}$ and in the brain parenchyma $\Omega_{\rm PAR}$ via
a mixed-dimensional transport model~\cite{masri2024modelling,laurino2019derivation} over a
timescale of minutes to hours (\Cref{fig:results1}A--C,
\Cref{fig:csf}A--C). Specifically, for $t > 0$, we solve for a
concentration $c = c(x, t)$ in the 3D intracranial compartments ($x
\in \Omega_{\rm CSF}, x \in \Omega_{\rm PAR}$) and for a cross-section
averaged concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial
and perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$) such that
the following equations hold:
\begin{subequations}
\begin{alignat}{2}
  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma & = 0 && \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}},
  \label{eq:multi_transport_3d}
  \\ 
  \partial_t (A  \hat c) - \partial_s(\hat D A \partial_s ( \hat c)) +\partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c})  &= 0 && \quad \quad \mathrm{in} \quad  \Lambda_a, \Lambda_v .
  \label{eq:multi_transport_1d}
 \end{alignat}%
\label{eq:multi_transport}%
\end{subequations}%
In~\eqref{eq:multi_transport}, $0 < \phi \leqslant 1$ is the fluid
volume fraction, $D$ and $\hat{D}$ are effective diffusion
coefficients, and $\bm u$ and $\hat{u}$ are fluid (CSF/ISF)
velocities, all defined in the respective compartments; $P = 2 \pi
R_2$ is the perimeter and $A = \pi (R_2^2 - R_1^2)$ the area of the
PVS cross-sections, $\delta_{\Gamma}$ is a Dirac term concentrated on
the interfaces $\Gamma_a, \Gamma_v$, the overlined $\overline{c}$
denotes the average of $c$ over cross-sections of $\Gamma_a,
\Gamma_v$, and $\xi$ is a permeability allowing for transfer/exchange
over the interfaces $\Gamma_a, \Gamma_v$ between the periarterial and
perivenous networks and their surroundings.  Moreover, we model the
membrane between the parenchyma and the CSF spaces as a semi-permeable
membrane with permeability coefficient $\beta_{\rm pia}$. For a
complete description of this mathematical model and simulation
scenarios, see \Cref{sec:app:more_maths}. The initial concentrations
in the three-dimensional domain $\Omega$ and networks $\Lambda_a,
\Lambda_v$ are all set to zero.

\subsection*{Intracranial influx after intrathecal injection}
To represent molecular influx into the intracranial compartments
following intrathecal injection, without modelling the spinal
compartment explicitly, we prescribe a time-dependent molecular influx
over the interface towards the spinal CSF compartment with the condition
\begin{equation*}
  (D \nabla c - \bm u c ) \cdot \bm{n} = g_{\mathrm{influx}},  \quad \mathrm{on}  \quad \Gamma_{\mathrm{SSAS}}.
\end{equation*}
Here we set
\begin{equation*}
  g_\mathrm{influx}(t) = \frac{m_{\rm tot}}{T_{\rm max}^2 |\Gamma_{\rm SSAS}|} \max(0, T_{\rm max} - |t - T_{\rm max}|), 
\end{equation*}
which expresses a hat-shaped influx function with a peak at $T_{\rm max}$ and a total tracer injection of $m_{\rm tot}$ (\Cref{tab:parameters}).

\subsection*{Molecular clearance from the intracranial space and perivascular network ends}

We model molecular clearance, into the meningeal lymphatics or other
pathways, across the upper, outer (arachnoid) boundary
(\Cref{fig:results1}A, \Cref{fig:csf}C), proportional to the
concentration in the SAS with a rate constant $\beta_{\rm exit}$. We
set a no-flux condition at the end nodes of the perivascular networks,
not allowing for direct molecular clearance there.

\subsection*{CSF flow in the SAS and ventricular system}
%\label{sec:csf_fluid_vel}

We model CSF as an incompressible Newtonian fluid at low Reynolds and
Womersley numbers via the Stokes equations with viscosity $\mu$, and
account for the convective contribution of the flow induced by CSF
production in the choroid plexus. Specifically, we compute the
velocity $\bm u_{\rm CSF}$ and associated pressure $p_{\rm CSF}$ in
the SAS and ventricular system induced by a constant production at a
rate of $u_{\rm in}$ across the lateral ventricle walls and with
efflux across the upper, outer (arachnoid) boundary with efflux
resistance $R_0$; and then set $\bm u = \bm u_{\rm CSF}$ in
$\Omega_{\rm CSF}$ in~\eqref{eq:multi_transport}
(\Cref{tab:parameters}, \Cref{fig:csf}A--E, see also
\Cref{sec:app:more_maths}). We do not model bulk flow within the brain
parenchyma, except within the perivascular network segments extending
below the pial surface, and thus set $\bm u = 0$ in $\Omega_{\rm PAR}$
in~\eqref{eq:multi_transport}.

\subsection*{Dispersion in the CSF spaces}
%\label{sec:csf_dispersion}
%% MER: I'm commenting out labels as long as there is no numbering

The pulsatile flow of CSF in the SAS and ventricular system,
associated with the cardiac and respiratory cycles, substantially
enhances molecular diffusion through
dispersion~\cite{taylor1953dispersion, watson1983diffusion,
  asgari2016glymphatic, sharp2019dispersion, ray2021quantitative,
  troyetsky2021dispersion}. To account for these dispersive effects,
while bridging from the pulsatile flow time scale of seconds to a
molecular transport time scale of hours, we compute cardiac and
respiratory dispersion factors $R_c$ and $R_r$, both
spatially-varying, in $\Omega_{\rm CSF}$ (\Cref{fig:csf}A) by
combining computational estimates of the respective peak CSF pressure
gradients with theoretical estimates for shear-augmented (Taylor)
dispersion \cite{taylor1953dispersion, watson1983diffusion,
  sharp2019dispersion} (see \Cref{sec:app:dispersion} for a complete
description). Combining these effects, we then set $R = R_c + R_r$ and
define the effective diffusion coefficient $D = (1 + R) D^{\rm Gad}$
in $\Omega_{\rm CSF}$ in~\eqref{eq:multi_transport} in the baseline model.

\subsection*{Perivascular fluid flow induced by CSF pressure differences}

To estimate the contribution also to perivascular fluid flow from CSF
production, we impose the fluid pressure $p_{\rm CSF}$ induced by CSF
production, computed in the SAS and ventricular system, at the end
nodes of the periarterial network located within the SAS. For the
network end nodes located within the parenchyma, we compute a harmonic
extension of the CSF pressure field (by solving a Laplace equation in
$\Omega_{\rm PAR}$ with $p_{\rm CSF}$ as the boundary value), and set
the corresponding value at these end nodes. We then numerically solve
a system of hydraulic network equations to compute $\hat{u}^p_a$
defined over $\Lambda_a$; i.e., the pressure-difference induced CSF
flow velocity in the periarterial network (see
also~\Cref{sec:app:hydraulic_network}). We repeat this procedure to
compute a corresponding velocity $\hat{u}^p_v$ in the perivenous
network $\Lambda_v$.

%% \subsection*{Dispersion in the PVS}

%% Similar to the CSF spaces, dispersion due to cardiac and respiratory pulsation has been hypothesized to enhance transport in the perivascular spaces. However, recent estimates indicate that the effect of dispersion does not exceed a factor of 2 compared to diffusion alone \cite{asgari2016glymphatic,sharp2019dispersion,bojarskaite2023sleep,asgari2016glymphatic,troyetsky2021dispersion}.

% \cite{sharp2019dispersion}: dispersion in PVS is negligiable in porous PVS, but might be relevant in open PVS (R=220) (theoretical estimates)
%  \cite{bojarskaite2023sleep} numerical simulations show that the enhancement factor is likely in the range of 10 - 50 \%
%\cite{asgari2016glymphatic} numerical simulations show increase of 16% to 50%  in distribution lengths
%\cite{troyetsky2021dispersion}:   maximum enhancement of diffusion of only about 5%

\subsection*{Perivascular fluid flow induced by arterial wall motion}

Using the theoretical framework introduced by Gjerde et
al.~\cite{gjerde2023directional}, we also compute analytic estimates
for the time-average perivascular flow rate $\langle Q' \rangle$
induced by peristalsis in the arterial network $\Lambda_a$; i.e.~the
net flow induced by traveling waves of arterial wall motion of
frequency $f$, amplitude $\varepsilon$ and wave lengths $\lambda$
(\Cref{fig:pvs}D). The corresponding contribution to the periarterial
flow velocity is defined for each segment $\Lambda_i$ as $\langle Q'_i
\rangle/A_i$ where $A_i = \pi (R_2^i - R_1^i)$ is the cross-section
area of the PVS segment and $\langle Q_i' \rangle$ is the estimated
mean flow rate (see Supplementary Methods~\ref{sec:app:peristalsis}
for more details). Two wall motion patterns are considered, yielding
$\hat{u}_a^c$ and $\hat{u}_a^v$ respectively: \emph{cardiac}
pulsations with $f = 1.0$Hz, $\lambda = 2$ m and $\varepsilon = 1\%$
(\Cref{fig:pvs}E), and very low-frequency \emph{vasomotion} with $f =
0.1$Hz, $\lambda = 0.02$m, and $\varepsilon = 10\%$
(\Cref{fig:pvs}F). No flow induced by peristalsis is considered for
the perivenous network. In the baseline model, we combine the
contributions from CSF production and cardiac peristalsis to set
$\hat{u}$ in~\eqref{eq:multi_transport} as $\hat{u}_a = \hat{u}_a^p +
\hat{u}_a^c$ in $\Lambda_a$, and $\hat{u}_v = \hat{u}_v^p$ in
$\Lambda_v$.

\subsection*{Model and material parameters}

Model and material parameters are summarized in \Cref{tab:parameters},
and~\Cref{tab:scenarios} gives an overview of the different in-silico
scenarios. We assume that the porosity equals the extracellular space
volume fraction $\phi = \phi_{\rm PAR}$ within the parenchyma and
$\phi = 1$ elsewhere. The effective diffusion coefficient is set equal
to that of Gadubutrol in free water in the PVS networks: $\hat{D} =
D^{\rm Gad}$; weighted by a dispersion factor $R$ in the CSF spaces:
$D|_{\Omega_{\rm CSF}} = ( 1 + R) D^{\rm Gad}$; and modulated by the
porosity and tortuosity ($\lambda = 1.78$) in the parenchyma:
$D|_{\Omega_{\rm PAR}} = D^{\rm Gad}_{\rm PAR}$. We set $\xi$
depending on whether the PVS segment is fully embedded within the CSF
spaces ($\xi = \xi_{\rm CSF}$) or parenchyma ($\xi = \xi_{\rm EF}$),
or neither (\Cref{sec:app:more_maths}). Additionally, we increase the
permeability $\xi$ ($100\times$) \fixme{near} the leaf nodes of
$\Lambda_a$ to account for the transition from explicitly resolved
PVSs to their unresolved continuation in the CSF spaces and
parenchyma.
\begin{table}
  \begin{center}
    \begin{tabular}{ll|ccc}
      \toprule
      Parameter& Symbol & Value & Unit& Reference\\
      \midrule
      Extracellular space volume fraction & $\phi_{\rm PAR}$ & 0.2 & -- & \cite{nicholson1981ion} \\
      Free diffusion coefficient, Gadubutrol & $D^{\rm Gad}$ & $3.8 \times 10^{-4}$& $\text{mm}^2/\text{s}$ & \cite{valnes2020apparent}\\
      Parenchymal diffusion coefficient, Gadubutrol & $D^{\rm Gad}_{\rm PAR}$ & $1.2 \times 10^{-4}$ & $\text{mm}^2/\text{s}$  & \cite{hornkjol2022csf} \\
      Permeability coefficient, pia & $\beta_{\rm pia}$ & $2.6 \cdot 10^{-8}$ & m/s & \cite{riseth2025twocompartment} \\
      Permeability coefficient, endfoot sheath & $\xi_{\rm EF}$ & $3.8\cdot 10^{-7}$  & m/s & \cite{koch2023estimates} \\
      Molecular efflux rate & $\beta_{\rm exit}$ & $10^{-4}$ & mm$^2$/s & \cite{hornkjol2022csf} \\
      Total molecular volume & $m_{\rm tot}$ & $0.5$ & mmol & \cite{eide2024functional} \\
      Molecular influx, time-to-peak & $T_{\rm max}$ & $1$ & h & \fixme{$\ast$} \\
      CSF production rate & $u_{\mathrm{in}}$ & $0.63$  & l / day & \cite{nilsson1992circadian} \\
      CSF viscosity & $\mu$ & $0.7$ & mPa$ \cdot $s & \cite{bloomfield1998effects} \\ 
      CSF outflow resistance & $R_0$ & $10^{-5}$  & Pa/(mm$\cdot$s) & \cite{hornkjol2022csf} \\ 
      Dispersion factor, CSF spaces & $R$ & $\dagger$ & -- & $\ast$ \\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{Model parameters. $\dagger$/$\ast$ denotes computed/estimated within this work.}
  \label{tab:parameters}
\end{table}
\begin{table}
  \begin{center}
  \begin{tabular}{ll|llll|lll}
    \toprule
    & Description & $R_1, R_2$ & $\xi_{\rm PVS-CSF}$ & $D|_{\Omega_{\rm CSF}}$ & $D|_{\Omega_{\rm PAR}}$ & $\mathbf{u}|_{\Omega_{\rm CSF}}$ & $\hat{u}|_{\Lambda_a}$ & $\hat{u}|_{\Lambda_v}$  \\
    \midrule
     & baseline & $R_2 = 2 R_1$ & $\xi$ & $(1 + R) D^{\rm Gad}$ & $D^{\rm Gad}_{\rm PAR}$ & $\mathbf{u}_{\rm CSF}$ & $\hat{u}_a^p + \hat{u}_a^c$  & $\hat{u}_v^p$ \\ 
    & no CSF production & $-$ & $-$ & $-$ & $-$ & $\mathbf{0}$ & $\hat{u}_a^c$ & $0$ \\ 
    & reduced pulsatility & $-$ & $-$ & $ (1 + 0.1 R) D^{\rm Gad}$ & $-$ & $-$ & $-$  & $-$ \\ 
    & increased CSF dispersion & $-$ & $-$ & $ (1 + 10 R) D^{\rm Gad}$ & $-$ & $-$ & $-$  & $-$ \\ 
    & higher PVS flow & $-$ & $-$ & $-$ & $-$ & $-$ & $\hat{u}_a^p + \hat{u}_a^c + \hat{u}_a^v$  & $-$ \\ 
    & higher PVS exchange & $-$ & 100$\xi$  &  $-$ & $-$ & $-$ & $-$  & $-$ \\ 
    & dilated PVS & $R_2 = 3 R_1$ & $-$ & $-$ & $-$ & $-$ & $\tilde{u}_a^p + \tilde{u}_a^c$  & $-$ \\ 
    \bottomrule
    \end{tabular}
    \end{center}
  \caption{Overview of models and model variations. ($-$ denotes equal
    to the corresponding baseline value). Each PVS cross-section is
    modelled as a concentric annulus with inner radius $R_1$ and outer
    radius $R_2$ and such that $R_2 = \beta R_1$. Let $\xi$ denote the
    astrocytic endfeet permeability estimated by Koch et
    al.~\cite{koch2023estimates}. The diffusion coefficients $D^{\rm
      Gad}$ represent the (free) diffusion coefficient of Gadubutrol
    in CSF (water), and $D^{\rm Gad}_{\rm PAR}$ its effective
    diffusion coefficient in human cortical tissue, all at body
    temperature~\cite{sykova2008diffusion, valnes2020apparent}. $R$ is
    the combined cardiac and respiratory dispersion enhancement field
    (see Methods). $\mathbf{u}_{\rm CSF}$ is the CSF velocity field
    induced by CSF production in $\Omega_{\rm CSF}$, and $\hat{u}_a^p$
    and $\hat{u}_v^p$ the axial velocity induced in the periarterial
    and perivenous networks, respectively, by the corresponding
    pressure differences. $\hat{u}_a^c$ and $\hat{u}_a^v$ is the
    periarterial velocity induced by arterial pulse wave wall motion
    and slow vasomotion, respectively. Last, $\tilde{u}_a^c$ and
    $\tilde{u}_a^p$ are as $\hat{u}_a^c$ and $\hat{u}_a^p$,
    respectively, but with dilated PVS.}
\label{tab:scenarios}
\end{table}

\subsection*{Numerical approximation, implementation, verification}

A complete description of the numerical methods used is provided in
\Cref{sec:details_numerical_method}, while a condensed summary is
given here. To solve~\eqref{eq:multi_transport} numerically, we use
the interior penalty discontinuous Galerkin (DG) method with weighted
averages and upwinding for the convection term in the 3D
domains~\cite{ern2009discontinuous} -- to ensure numerical stability
and minimize the artificial presence of negative concentration
values. For the 1D domain, we use continuous finite element spaces
defined over each of the networks and stabilized with numerical
diffusion when $\hat{u}$ is non--zero. Note that the 3D--1D coupling
in~\eqref{eq:multi_transport} leads to a 3D solution $c$ with low regularity properties. Thus, one can only expect lower-order error convergence near
$\Lambda_a$ and $\Lambda_v$, see \cite{masri2024discontinuous} for the numerical analysis of DG schemes and \cite{laurino2019derivation} for continuous Galerkin schemes.  However, away from these 1D networks, almost
optimal approximation is expected~\cite{masri2023discontinuous,koppl2016local}. To numerically solve the Stokes equations to
compute the CSF flow velocity and pressure in the ventricular system
and SAS, we use finite element spaces that preserve the
incompressibility condition on the discrete
level~\cite{hong2016robust}. This requirement is important for
numerical stability when later
solving~\eqref{eq:multi_transport} %\cite{cesmelioglu2022compatible}.
 The
spaces we used for the velocity are continuous along the normal
direction; continuity of the tangential component is enforced via
interior penalty approaches~\cite{hong2016robust}. Optimal
approximation properties are expected for this scheme. These numerical
methods were implemented using the FEniCS finite element
framework~\cite{alnaes2015fenics}, and the 3D-1D coupling is handled
by the extension
FEniCS\textsubscript{ii}~\cite{kuchta2020assembly}. The correctness
and accuracy of the numerical solutions were verified by a series of
numerical verification experiments \fixme{(see \Cref{sec:app:verification})}.

%% The choice this scheme was is necessary here since continuous Galerkin
%% methods even when supplemented with artificial diffusion stabilization
%% terms were unstable in our experiments.

\subsection*{Model validation}

As the primary means of model validation, we compare the in-silico
predictions of tracer enrichment and clearance against glymphatic MRI
studies (Discussion). In addition, we compare auxiliary model
quantities (CSF flow rates and pressure differences, dispersion factor
estimates, and the shapes and sizes of PVSs) with the current
literature (Supplementary discussion: Extended model validation).

\section*{Acknowledgments}

All authors express their gratitude to Prof.~Jürgen Reichenbach for
generously sharing the multi-modal human MR data, and to Prof. Erlend
Hodneland for facilitating this data exchange. M.E.R. acknowledges the
support of the Foundation Kristian Gerhard Jebsen via the K.~G.~Jebsen
Center for Brain Fluid Research and the Research Council of Norway via
grant \#324239 (EMIx). Simulations were performed on the Experimental
Infrastructure for Exploration of Exascale Computing (eX3), which is
financially supported by the Research Council of Norway under contract
270053.

\section*{Author contributions statement}

M.C., M.K, R.M., and M.E.R. conceived and designed the project.
M.C., M.K, R.M., and M.E.R. contributed to software development. 
M.C., M.K, and R.M. conducted the experiments.
M.C. analyzed the results and prepared the figures.
M.C. M.K. and R.M and M.E.R wrote the manuscript. 
All authors edited and reviewed the manuscript.

\section*{Competing interests}

The authors declare no competing interests.

\bibliography{references}

\newpage
\FloatBarrier
\appendix
\input{supplementary_main.tex}

%\newpage
%\FloatBarrier
%\input{background.tex}

\end{document}